---
title: "HW04 Part 1: Complete Chapter 3"
author: "Dylan Klingeman"
date: "`r format(Sys.time(), '%d %B %Y')`"
output: html_notebook
editor_options: 
  chunk_output_type: inline
---



```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Chapter 3

- Change "your name" in the YAML header above to your name.

### Section 3.1

Load the tidyverse packages. I've entered the first code chunk for you. **You must enter all subsequent code chunks and run the code.** 

```{r}
library("tidyverse")
```

If you get an error, then you did not install the `tidyverse` package like you were supposed to in the previous assignment. Go back to that assignment and complete that step. (This should not be the case, though.)

### Section 3.2: First steps

Enter your code chunks for Section 3.2 here.

# should display a chart comparing displ adn hwy
```{r}
ggplot(data = mpg) + 
    geom_point(mapping = aes(x = displ, y = hwy))
```

# should create a graph of data, geom function, and mappings when replaced

```{r}
ggplot(data = DATA) + 
  GEOM_FUNCTION(mapping = aes(MAPPINGS))
```


Get in the habit of entering a brief description of what each chunk of code does. Enter this above the chunk, like I did in Section 3.1.

#### Section 3.2 Questions

Answer the questions *completely.* Some answers will require you to write code in a code chunk and also to type text before or after the chunk. Other answers will require only code or only text. 

**1:** Run ggplot(data = mpg). What do you see?

```{r}
ggplot(data = mpg)
```

Running the above function displays a blank text box. Doing the following instead gives us a completed graph with data.On engine size and fuel efficiency.  

```{r}
ggplot(data = mpg) + 
    geom_point(mapping = aes(x = displ, y = hwy))
```

#

**2:** How many rows are in `mpg`? How many columns? **Hint:** Use the `dim()` function. Type `dim()` in the console to learn what `dim()` does.

```{r}
dim(mpg)
```

The dim function shows us that there are 234 rows and 11 columns within the mpg data set. 

#

**3:** What does the `drv` variable describe? Read the help for `?mpg` to find out.

"Drv" refers to the type of "drive train" which is how the car operates. For example the variable "f" refers to a front wheel drive, while "4" refers to a four wheel drive. 

#

**4:** Make a scatterplot of `hwy` vs `cyl`.

Below is the scatterplot for hwy vs cyl

```{r}
ggplot(data = mpg) + 
    geom_point(mapping = aes(x = hwy, y = cyl))
```

# 

**5:** What happens if you make a scatterplot of `class` vs `drv`? Why is the plot not useful?

```{r}
ggplot(data = mpg) + 
    geom_point(mapping = aes(x = class, y = drv))
```

Plotting these two variables against each other shows the type of driving engines in each of the classes of cars. This could be viewed i other more efficient ways within the data set. No new information is really being compared here. 

#

### Section 3.3: Aesthetic mappings

shows the plot of displ vs hwy, but also plots the class into different color based categories. 
```{r}
ggplot(data = mpg) + 
  geom_point(mapping = aes(x = displ, y = hwy, color = class))
```


shows the plot of displ vs hwy, but also plots the class itno different sizes depending on the category

```{r}
ggplot( data = mpg) + 
    geom_point(mapping = aes(x = displ, y = hwy, size = class))
```


shows the plot of displ vs hwy, but also plots the class into different transparencies by category

```{r}
ggplot(data = mpg) + 
    geom_point(mapping = aes(x = displ, y = hwy, alpha = class))
```


shows the plot of displ vs hwy, but also plots the class into different shapes by category

```{r}
ggplot(data = mpg) + 
    geom_point(mapping = aes(x = displ, y = hwy, shape = class))
```

#

#### Section 3.3 questions

**1:** What’s gone wrong with this code? Why are the points not blue?

Below is the code block used in the question: 

```{r}
ggplot(data = mpg) + 
  geom_point(mapping = aes(x = displ, y = hwy, color = "blue"))
```

THe color function is not working since it is within the same set of parantheses as x and y, meaning it is being compared to those rather than working on. The proper code block would look like the following: 

```{r}
ggplot(data = mpg) + 
    geom_point(mapping = aes(x = displ, y = hwy), color = "blue")
```

# 

**2:** Which variables in `mpg` are categorical? Which variables are continuous? (*Hint:* type `?mpg` to read the documentation for the dataset). How can you see this information when you run mpg?


Categorical: "manufacturer", "model", "trans", "drv", "class",  "fl"

Continuous: "displ", "year", "cyl", "cty", "hwy"

When the table is ran there is a small <> beneath it that describes the classification

#

**3:** Map a continuous variable to color, size, and shape. How do these aesthetics behave differently for categorical vs. continuous variables? **Note:** You need several code chunks here to answer this question.

# Color

Mapping a continuous variable like "cyl" elow will order the colors into a gradient. 

```{r}
ggplot(data = mpg) + 
  geom_point(mapping = aes(x = displ, y = hwy, color = cyl))
```

Mapping a categorical variable like "manufacturer" will sort hte data into many different colors. 

```{r}
ggplot(data = mpg) + 
  geom_point(mapping = aes(x = displ, y = hwy, color = manufacturer))
```

# Size

Mapping the continuous variable like "cty" will change the sizes of different points based on the city. This works correctly. 

```{r}
ggplot(data = mpg) + 
  geom_point(mapping = aes(x = displ, y = hwy, size = cty))
```

When using the "size = x" function, it will give an error message if a categorical/discrete variable is used. The data overlaps each other and is hard to interpret. 

```{r}
ggplot(data = mpg) + 
  geom_point(mapping = aes(x = displ, y = hwy, size = class))
```

# Shape

It is not possible to use teh shape function on a continuous variable. As shown below it will report an error and not complete the graph.

```{r}
ggplot(data = mpg) + 
  geom_point(mapping = aes(x = displ, y = hwy, shape = year))
```

Mapping a categorical variable works for the shapes, but having this many in the data set leads to errors with running out of shapes. 

```{r}
ggplot(data = mpg) + 
  geom_point(mapping = aes(x = displ, y = hwy, shape = class))
```


**4:** What happens if you map the same variable to multiple aesthetics?

Having more than one type of aesthetic will cause them to overlap and show on the same thing, as shown below with "cty". THere is both a color gradient and a change in size depending on the data point. 

```{r}
ggplot(data = mpg) + 
  geom_point(mapping = aes(x = displ, y = hwy, size = cty, color = cty))
```


**5:** What does the stroke aesthetic do? What shapes does it work with? (*Hint:* use `?geom_point`)

The stroke aesthetic seems to greatly enlarge the size of certain data points, including other aesthetics applied. 

```{r}
ggplot(data = mpg) + 
  geom_point(mapping = aes(x = displ, y = hwy, stroke = cty, shape = trans))
```


**6:** What happens if you map an aesthetic to something other than a variable name, like `aes(colour = displ < 5)`?

Using a function like the one listed above will apply a color (blue), when the condition is met that it is set to. In this case being below 5 will color the variable blue. 

```{r}
ggplot(data = mpg) + 
  geom_point(mapping = aes(x = displ, y = hwy, color = displ < 5))
```


### Section 3.5: Facets



#### Section 3.5 questions

**1:** What happens if you facet on a continuous variable? **Note:** Write a prediction, then make and run your code chunk.



**2:** What do the empty cells in plot with `facet_grid(drv ~ cyl)` mean? How do they relate to this plot?



**3:** What plots does the following code make? What does `.` do?



**4:** Take the first faceted plot below. What are the advantages to using faceting instead of the colour aesthetic? What are the disadvantages? How might the balance change if you had a larger dataset?



**5:** Read `?facet_wrap`. What does `nrow` do? What does `ncol` do? What other options control the layout of the individual panels? Why doesn’t `facet_grid()` have nrow and ncol argument?


**6:** When using `facet_grid()` you should usually put the variable with more unique levels in the columns. Why?

### Section 3.6: Geometric objects



#### Section 3.6.1 Questions

**1:** What geom would you use to draw a line chart? A boxplot? A histogram? An area chart? **Note:** You will have to make some types of these charts in Part 2 of this assignment.


**2:** Run this code in your head and predict what the output will look like. Then, run the code in R and check your predictions.



**3:** What does `show.legend = FALSE` do? What happens if you remove it?
**Note:** Skip the "Why do you think I used it earlier in the chapter?" question.


**4:** What does the `se` argument to `geom_smooth()` do?


**5:** Will these two graphs look different? Why/why not?


**6:** Recreate the R code necessary to generate the following graphs. **Note:** Enter a separate code chunk for each graph, numbered 1 to 6, starting row-wise at upper left. 

1: upper left.


2: upper right.


3: middle left.


4: middle right.


5: lower left.


6: lower right. *Hint:* Read the help file for `geom_point`. Use `shape = 21`, `stroke = 2`, and `size = 3`.

### Section 3.7: Statistical transformations.

**Skip this section.** We may return to it later.


### Section 3.8 Position adjustments


#### Section 3.8 Questions

**1:** What is the problem with this plot? How could you improve it. *Hint:* Remember the size of the `mpg` data set from earlier in this exercise. Does the number of points seem correct? **Note:** Write the code to improve it. 


**2:** What parameters to `geom_jitter()` control the amount of jittering?


**3:** Compare and contracts `geom_jitter()` with `geom_count()`.


**4:** What is the default position adjustement for `geom_boxplot()`? Create a visualization of the `mpg` dataset that demonstrates it.


### Section 3.9: Coordinate systems

**Note:** Read about and do only the code chunk associated with `coord_flip()`. `coord_flip()` is a great layer to use when you have categorical data with longish names as it makes the text easier to read.

### Section 3.10: The layered grammar of graphics

Just read this section as a contextual review of everything you just learned.

