---
title: "HW 12: Dates"
author: "Dylan Klingeman"
date: "`r format(Sys.time(), '%d %B %Y')`"
output: html_notebook
editor_options: 
  chunk_output_type: inline
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

***
## Libraries
```{r}
library(tidyverse)
library(here)
library(lubridate)
```

```{r}
source(here("scripts", "my_functions.R"))
```

### Part 1: Play Date

```{r}
birth_date <- "04/03/1996"

# Elizabeth Garret Anderson was the first female physician
anderson <- "9 June 1836"

# Mary Anning discovered the first complete Plesiosaur fossil, 
# and other important Jurassic fossils
anning <-  "1799/05/21"

# Alice Augusta Ball was a chemist who developed the
# first injectable treatment for leprosy.
ball <- "July 24, 1892"

# Sylvia Earle was the first chieft scientist at NOAA
earle <- "August, 30 (1935)"

# Rosalind Franklin used x-ray diffraction to 
# discover DNA structure. Watson and Crick claimed credit.
franklin <- "25th of July, 1920"

# Ester Lederberg was the first to discover that viruses 
# infect bacteria. Led to discovery of lambda bacteriophage
lederberg <- "December 18, 1922"

# Barbara McClintock discovered transposable elements (jumping genes),
# and developed staining techniques to study chromosomes
mcclintock <- "16061902"

# Let's also remember a dream
mlk_birth <- "1/15/29"
mlk_dream <- "1963 August 28"
```

## 1.1

```{r}
my_birth_date <- mdy(birth_date)
my_birth_date

anderson_birth_date <- dmy(anderson)
anderson_birth_date

anning_birth_date <- ymd(anning)
anning_birth_date

ball_birth_date <- mdy(ball)
ball_birth_date

earle_birth_date <- mdy(earle)
earle_birth_date

franklin_birth_date <- dmy(franklin)
franklin_birth_date

lederberg_birth_date <- mdy(lederberg)
lederberg_birth_date

mcclintock_birth_date <- dmy(mcclintock)
mcclintock_birth_date

mlk_birth_date <- mdy("01/15/1929")
mlk_birth_date

mlk_dream_date <- ymd(mlk_dream)
mlk_dream_date
```

# This code chunk inserts all of the birthdays from some important figures, as well as the i have a dream speech. 

#

## 1.2

# Fixed the MLK birth date above. 

#

## 1.3
```{r}
wday(my_birth_date, label = TRUE, abbr = FALSE)
```

#

## 1.4

```{r}
updated_my_birthday <- update(my_birth_date, year=2020)
weekdays(updated_my_birthday)
```

# I cannot find this one even by googling it, will come back to it. 

## 1.5 

```{r}
updated_my_birthday <- update(my_birth_date, year=2525, month = 4, day = 03)
wday(updated_my_birthday, label = TRUE, abbr = TRUE)

year(updated_my_birthday) - year(my_birth_date)

```

# 

## 1.6

```{r}
day(ball_birth_date) - day(anderson_birth_date) 
```

# This code chunk subtracts the birthdates between ball and anderson. 

#

## 1.7

```{r}
(mlk_dream_date - mlk_birth_date)
```

# 1.8 

```{r}
birth_dates <- c(my_birth_date, anderson_birth_date, anning_birth_date, ball_birth_date, earle_birth_date, franklin_birth_date, lederberg_birth_date, mcclintock_birth_date, mlk_birth_date)
names(dates) <- c('Klingeman', 'Anderson', 'Anning', 'Ball', 'Earle', 'Franklin', 'Lederberg', 'McClintock', 'King')

leap_year(birth_dates)

birth_dates - my_birth_date
```

#

### Part 2: Wrangling and plotting with dates


## 2.1 Lake_Ice

```{r}
ice_data <- read_csv(here::here("data","lake_ice.csv"),na="-999") %>% 
drop_na()
ice_data                 
```

# Imports lake_ice.csv dataset to be used. Also converts any -999 to NA values, then removes the NA values. If this 999 was left in it would heavily influence the graphs later on. 


#

```{r}
wrangled_ice <- filter(ice_data, lakename=="LAKE MENDOTA")

select(wrangled_ice, iceon_year, iceon_month, iceon_day, iceoff_year, iceoff_month, iceoff_day) %>% 
  mutate(ice_on = make_date(iceon_year, iceon_month, iceon_day)) %>% 
  mutate(ice_off = make_date(iceoff_year, iceoff_month, iceoff_day)) %>% 
  mutate(duration = as.numeric(ice_off - ice_on))

```

# This code chunk wrangles the data by filtering only for lake mendota within the data set. Then it is filtered to only show the data on the iceon/off days/months/years, which are then combined into single columns. These columns (ice_on and ice_off) are then subtracted from each other to create another final column titled duration. Duration measures the amount of tiem that ice was present in days. 

#

```{r}
wrangled_ice %>% 
  ggplot() + 
  geom_line(aes(x = iceoff_year, y = duration), na.rm = TRUE) + 
  labs(x = "Year", y = "Ice Duration (days)")
```

# This code chunk creates the graph for the Lake ice dataset. This is almost identical to the one shown, but I noticed my value around the year 2000 is to the right of the line instead of to the left. 
# Removed console error by moving () from behind TRUE to behind duration. 

#

## 2.2 Soil Carbon Dioxide. 

```{r}
soil_data <- read_tsv(here::here("data","soil_co2.txt"))
```

# Imports the soil_co2 dataset. 

#

```{r}
soil_date <- dmy(01012006)
soil_date
```

# Adds the date 01/01/2006 to a vector. 

#

```{r}
renamed_soil <- soil_data %>% 
  rename("day" = "DayN[]", "datecode" = "Datecode[]", "min" = "Min[]", "veg" = "Veg[]", "plot" = "Plot[]", "group" = "Group[]", "corrco2" = "CorrCO2[]", "soilt" = "SoilT[]") %>% 
  mutate(date = soil_date + day, 
         year = year(date), 
         month = month(date, label =TRUE))
renamed_soil
```

# This code chunk wrangles the soil_data datset. The names are changed to a shorthand that are more coding friendly, then adds three new columns: date, year, and month. 

#

```{r}
wrangled_soil <- group_by(renamed_soil, year, month)
wrangled_soil
```

# Organizes data by the month and year

#

```{r}
soil_summary <- summarise(wrangled_soil, meanco2=mean(corrco2, na.rm = TRUE),std_err_co2 = std_err(corrco2, na.rm = TRUE), .groups = 'drop')  
soil_summary
```

# Summarises the data into only year, month, the std error of the mean, and the mean co2 value. 

#

```{r}
ggplot(data = soil_summary) + 
  geom_col(mapping = aes(x = month, y = meanco2)) + 
  geom_errorbar(aes(x = month, ymin = meanco2 - std_err_co2, ymax = meanco2 + std_err_co2), width = 0.65) + 
  facet_wrap(~year) + 
  labs (x = "Month", y = "mean_co2")
```

# Creates the first set of charts as seen in the example. I opted to use the errrorbar and set it to 0.65 width, this seemed to be easy to see but not invasive of other points. 

# The graph was originally showing numbers instead of the months. Fixed this by goign back a few steps and adding label = TRUE argument to the months. 

#

```{r}
ggplot(data = soil_summary) + 
  geom_point(mapping = aes(x = month, y = meanco2)) + 
  geom_linerange(aes(x = month, ymin = meanco2 - std_err_co2, ymax = meanco2 + std_err_co2)) + 
  facet_wrap(~year) + 
  labs (x = "Month", y = "mean_co2")
```

# Makes a graph similar to the above but changes it to a line range graph with points instead of columns. 

#

## 2.3 Mississippi River Discharge. 

```{r}
thebes_data <- read_csv(here::here("data","thebes_discharge.csv"))
```

# Imports thebes_discharge.csv dataset

#

```{r}
wrangled_thebes <- pivot_longer(thebes_data, "1988":"2018", names_to = "year", values_to = "discharge") %>% 
  mutate(discharge = discharge/1000) %>% 
  group_by(year, Month)
wrangled_thebes
```

# Wrangles the dataset, collapses the columns from 1988 to 2018 into one column called "year" and these values are transferred to another column titled "discharge'. This value is then divided by 1000 for unit conversions. The data is then sorted by year and month. 

#

```{r}
thebes_summary <- summarise(wrangled_thebes, mean_discharge = mean(discharge, na.rm = TRUE),.groups = 'drop')
thebes_summary
```

# This code chunk calculates the mean discharge for each month of the dataset. 

#

```{r}
thebes_named <- thebes_summary %>% 
  mutate(event = case_when(
    year==1993 | year==2011 ~ "Flood", 
    year==1988 | year==2012 ~ "Drought", 
    TRUE ~ "Normal")) %>% 
    mutate(Month = month(Month, label = TRUE))
thebes_named
```

# This code chunk creates another column labeled as event. This tracks when there was a drought or when there was a flood changing the system. If there was not a drought or flood that year, then it is labeled as "normal". The TRUE statement is needed for the non specified years. 

#

```{r}
thebes_events <- thebes_named %>% 
  filter(year%in%c("1988", "1993", "2011", "2012"))
thebes_events
```

# Filters only years where there was a year or a drought. I was able to get this to work with "==" without the %in%, but after going back and looking at chapter 5.2.2 i was able to make the %in% argument work. 

#

```{r}
ggplot(data = thebes_named) +
geom_line(mapping = aes(x = Month, y = mean_discharge, group = year), size = 0.3, color = 'gray70') +
geom_line(data = thebes_events, mapping = aes( x = Month, y = mean_discharge, group = year, color = event)) + 
  labs(x = NULL, y = "Discharge(cf * s^-1, thousands)")
  
  
```

# Creates the graph for the thebes dataset. The month is plotted to x, while the discharge is plotted to y. The normal years are plotted in size 0.3 lines and have their color changed to gray70. The other drought/flood lines have their color changed and are larger. The axes have their names changed to better match the graph seen in the dataset. 

# The lines were not showing separately until i placed in group = year. 

#