---
title: "HW05"
author: "Dylan Klingeman"
date: "`r format(Sys.time(), '%d %B %Y')`"
output: html_notebook
editor_options: 
  chunk_output_type: inline
---


```{r}
library("tidyverse")
```


### Plot 1: Scatterplot of the trees data set

```{r}
data(trees)
?trees
```

acquires the data set for "trees' and examines the properties of the data set. 

```{r}
ggplot(data = trees, mapping = aes(x = Height, y =Girth)) +
geom_point()
```

creates a scatterplot comparing Height vs Girth from the trees data set. 

#


### Plot 2: Linegraph of the lynx dataset

```{r}
data(lynx)
?lynx
```

acquire the data set for "lynx" and examines the properties of the data set. 

```{r}
year <- c(1821:1934)
year
```

creates a vector named "year" which contains the numbers 1821-1934. The data within the vector is then displayed

```{r}
class(lynx)
```
displays the class of data that the "lynx" dataset is, which is a Time Series (ts)

```{r}
pelts =as.vector(lynx)
pelts
```

converts the lynx data set into a vector. 

```{r}
lynx_pelts <- data.frame(year, pelts)
lynx_pelts
```

Creates a dataframe comparing the vectors "year" and "pelts"

```{r}
ggplot(data=lynx_pelts) + 
  geom_line(mapping = aes(x = year, y = pelts),color="maroon")
```

Creates a line graph comparing the "year" vector to the "pelts" vector within the data set "lynx_pelts". The color of the line is then changed to maroon. 

#

### Plot 3: Scatterplot of the iris data set


```{r}
data(iris)
?iris
```

acquire the data for "iris" and examines the properties of the data set.

```{r}
ggplot(data = iris, mapping=aes(x = Petal.Length, y = Petal.Width)) + 
  geom_point(mapping = aes(color = Species, shape = Species, size=2)) +
  labs(x = "Petal Length", y = "Petal Width")
```

Creates a scatterplot of the iris data set that compares the "Petal.Length" to "Petal.Width". There is also a third variable, "Species", which is shown by the differences in color and shape. The size of the points is then changed to 2. 

#

### Plot 4 and 5: 

