---
title: "Useful Title"
author: "Dylan Klingeman"
date: "r format (Sys.time(), '%d %B %Y')'"
output: html_notebook
editor_options: 
  chunk_output_type: inline
---


```{r}
library(tidyverse)
library(knitr)
library(xtable)
```
```{r}
anscombe
```

# loads the libraries to be used later in the assignment

```{r}
ansc <- 
    bind_cols(
        anscombe %>%
            select(matches("x")) %>%
            gather(dataset, "x"), 
        anscombe %>%
            select(matches("y")) %>%
            gather(key, "y")
  ) %>%
  select(-key) %>%
           mutate(datasset = gsub("x", "", dataset)) %>%
           as_tibble()
    
```

# creates the dataset "ansc" 

```{r}
ansc
```

# should display the dataset created above

```{r}
ansc %>%
    group_by(dataset) %>%
    summarize(
      mean_x = mean(x), 
      stdev_x = sd(x), 
      mean_y = mean(y), 
      stdev_y = sd(y)
    )
```

# should displayy mean and standard deviation for both the x and y variables

```{r}
ansc %>% ggplot(aes(x, y, color = dataset)) + 
  geom_point() + 
  facet_wrap(~dataset, ncol = 2)
```

# should create a chart separanted into facets. the chart will have two columns as data due to "ncol = 2". 

```{r}
head(iris)
```

# should display the first rows/columns for the iris data set

```{r}
kable(head(iris), 
      allign= c("1", "c", "r", "c", "1"), 
      col.names = c("Sepal Length", "Sepal Width", "Sepal Length", "Sepal Width", "Species"))
```

# assigns placements for the columns in the iris data set


```{r}
kable(head(airquality), 
        align= c("r", "r", "c", "l", "l"), 
        col.names = c("Ozone (ppb)", "Solar R (lang)", "Wind (mph)", "Temp (F)", "Month", "Day"))
```

# assigns placements for the columns, as well as appropriate column names with units

```{r}
chicks_lm <- lm(weight ~ feed, data = chickwts)

summary(chicks_lm)
```

# loads statistical data of the chickwts data set, assigns it to "chicks_lm", then runs a summary of it in "summary(chicks_lm)

```{r}
kable(head(chicks_lm))

```

### Running kable on "chicks_lm" creates a table of the "chicks_lm" data, with x values for all of the seed types. The data seems all over the place and is difficult to interpret. 

```{r}
xtable(chicks_lm) %>% kable()
```

# produces an organized table for the chicks_lm dataset

```{r}
trees_lm <- lm(Girth ~ Height, data = trees)

summary(trees_lm)
```

# loads statistical data comparing hte girth and height of a tree, then lists out the dat in a summary with the "summary(trees_lm) argument. 

```{r}

xtable(trees_lm) %>% kable()
```

# organizes the data into a table, just like above in the xtable(chicks_lm) example

$p + \frac{1}{2}pV^2 + pgh = constant$ 

# creates a formula within rstudio 