---
title: "HW 09: More Data Wrangling II"
author: "Dylan Klingeman"
date: "`r format(Sys.time(), '%d %B %Y')`"
output: html_notebook
editor_options: 
  chunk_output_type: inline
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```
***
## Libraries
```{r}
library(tidyverse)
library(here)
library(knitr)
library(smatr)
```
***
```{r}
source(here("scripts", "my_functions.R"))
```


## Datasets
```{r}
darters_data <- read_csv(here("data", "darters.csv"))
```

#This code chunk loads in the darters.csv dataset from the data folder. 
***
## Wrangling Data
```{r}
wrangled_darters <- darters_data %>% 
  filter(species == "blennioides" | species == "caeruleum" | species =="spectabile"
  ) %>% 
  mutate(
      total_substrate = majsub + minsub
       ) %>% 
mutate(sex = ifelse(sex =="f", 
                           "female", 
                           "male")) %>% 
mutate(riffle = ifelse(riffle==1, 
                       "Riffle 1", 
                       "Riffle 2")) %>% 
mutate(length = length*10)

wrangled_darters
```

#This code chunk wrangles the filtered darers.csv dataset. The species "tetrazonum" and "zonale" are filtered out from the data set. A new column "total_substrate" is added, which adds majsub + minsub. The sex column is modified so "f" and "m" are changed to "female" and "male". The Riffle column is changed to list "Riffle 1" or "Riffle 2". The new data set is then listed. 

***

## Summary

```{r}
darter_summary_data <- group_by(wrangled_darters, species, riffle)
darter_summary <- summarise(darter_summary_data, 
                            mean_length = mean(length, na.rm = TRUE),
                            mean_velocity = mean(velocity, na.rm = TRUE),
                            mean_depth = mean(depth, na.rm = TRUE), 
                            .groups = 'drop')
darter_summary
```

#This code chunk was very difficult. First the wrangled_darters data froma bove is grouped by just the species and riffle. Then this is summarised and used to gather a mean value for: length, velocity, and depth. The issues were in having to look up the ".groups" argument and figuring out what it was interacting with the species. 

#error message from above: "`summarise()` regrouping output by 'species' (override with `.groups` argument)"

# Added na.rm = TRUE to remove NA values. (necessary for graph below)

## Kable

```{r}
kable(darter_summary)
```
#
***
## Graph Data
# Plot 1

```{r}
ggplot(data = darter_summary_data) + 
  stat_summary(mapping = aes(x = species, y = length), 
               fun.ymin = min,
               fun.ymax = max, 
               fun.y = median) +
  geom_point(data = darter_summary, mapping = aes(x = species, y = mean_length),
             shape = 15,
             color = 'firebrick') + 
  coord_flip() + 
  facet_wrap(~riffle) + 
  labs(y ="Length(in mm)", x = NULL)
```

#This code chunk creates a summary graph for the darter_summary from above. This shows the median, minimum, and maximum data points for each of the species. The mean+length is also shown as the colored square in red on each of the lines. A facet_wrap is also applied that separates the graph by the two riffles. The values are coord_flipped to better space out the names of the species. The y is relabeled from length to "Length (in mm).

# Riffle 1 did not have the mean length appear for blennioides, this was because NA values were not removed. 


Looking at each of the species we can see some large differences between the riffles. For the spectabile group, it has a very close minimum and maximum between riffles, but its median and mean are much higher in riffle 1 than riffle 2. 

The Caeruleum species is very similar, but the riffle 1 has a higher maximum of ~46, compraed to riffle 2's ~41. The mean for riffle 1 is also around ~36, compared to riffle 2's ~28. 

The blennioides had very similar median, minimum, and mean, but riffle 2 had a much larger maximum at ~56, compared to riffle 1's ~38. 

# 
***

# Plot 2

```{r}
ggplot(data = darter_summary_data) +
stat_summary(mapping = aes(x = species, y = length), 
             fun.y = median,
             fun.ymin = min, 
             fun.ymax = max, 
             geom = "point",
             size = 2.2, 
             na.rm = TRUE) + 
  stat_summary( mapping = aes(x = species, y = length),
               fun.y = mean, 
               fun.ymin = min, 
               fun.ymax = max, 
               size = 3, 
               color = "maroon", 
               na.rm = TRUE) + 
  facet_wrap(~riffle) + 
  coord_flip() + 
  labs(y = "Length(in mm)", x = NULL) +
geom_point(data = darter_summary, mapping = aes(x = species, y = mean_length), color = "grey60") 
```

# This code chunk creates a data summary comparing the length differences between 3 species in 2 different riffles. This graph uses teh same data as the one above, but features new aesthetic changes to improve clarity. Of note is the color argument and the point size argument. 

# geom_point has to be placed at the bottom to be visible on the graph. 

#

***
### Stress and corticosterones in birds

## Import Data

```{r}
feathers_data = read_csv(here("data", "feathers.csv"), locale = locale(decimal_mark = ","))
```

# Loads in the feathers_data dataset. The decimal marks are changed from '.' to ','. This is doen by the locale = locale() argument. 

## Rename Data

```{r}
renamed_feathers <- feathers_data %>% 
  rename(capture_date = "Capture Date", day = "Date", site = "Site", urban_score = "Urba Score", cortF = "CortF ng/mg", cortB = "CortF ng/mm", sex = "Sex", smi = "SMI", tarsus = "Tarsus", mass = "Mass", basal_cort = "Bas_Cort", stress_cort = "Stress_Cort")
renamed_feathers
```

# Renames the column names in the feathers_data set to make them easier to manage. 

#

## Wrangle Data

```{r}
mean_tarsus <- mean(tarsus)
```

# Takes the mean of the tarsus data set and creates a vector for it. 

```{r}
wrangled_feathers <- renamed_feathers %>% 
 mutate(log_mass = log(mass), 
        log_tarsus = log(tarsus))
wrangled_feathers
```

# Adds new columns for the logs of the mass values and tarsus values within the feathers dataset. 

```{r}
major_axis <- sma(log_mass ~ log_tarsus, data = wrangled_feathers)
```

# Calculates the standardized major axis regression for the data set.

```{r}
summary(major_axis)
```

# Summarizes the major axis, shows the slope value for estimate. THis needs to be extracted and saved as "b". 

```{r}
coef(major_axis)
slope <- coef(major_axis)
b <- slope['slope']
b
```

# extracts the slope of the estimate and stores it as the valeu "b". 

```{r}
mutated_feathers <- mutate(wrangled_feathers, new_smi=scaled_mass(mass,tarsus,b)) 

select(mutated_feathers, smi, new_smi)
```

# Creates a new column for the new scaled mass index. This  is then compared to the previous "smi" already in the dataset. The values are identical, with the new_smi having more decimals. 

#

## Summarize Data

```{r}
mutated_feathers_site <- group_by(mutated_feathers,site)
feathers_summary<- summarise(mutated_feathers_site,
    basal_mean =  mean(basal_cort, na.rm = TRUE),
    stress_mean = mean(stress_cort, na.rm = TRUE),
    std_err_basal=std_err(basal_cort, na.rm = TRUE),
    std_err_cort=std_err(stress_cort, na.rm = TRUE), 
    .groups = 'drop')
feathers_summary

```

# This code chunk begins by grouping the data by site in the mutated_fethers data set. This data is then summarised, while creating new columns for the "basal_mean", "stress_mean", and the standard error of those two. Removed .groups error with 'drop' argument. 

# Tried standard error on the basal_mean and stress_mean, changed it to basal_cort and stress_cort respectively. 

## Graph Data

# Plot 1

```{r}
ggplot(data = mutated_feathers) + 
  geom_boxplot(mapping = aes(x = site, y = new_smi)) + 
  coord_flip()
```

# Each site has some overlap within the dataset, but the median for each site differs slightly. The NIORT has a much lower minimum (around ~20), compared to the rest being above 22. The maximum value for each one is also very similar. 

# Plot 2