---
title: "HW 10: Data Visualization 1"
author: "Dylan Klingeman"
date: "`r format(Sys.time(), '%d %B %Y')`"
output: html_notebook
editor_options: 
  chunk_output_type: inline
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```
***
## Libraries
```{r}
library(tidyverse)
library(here)
library(GGally)
library(patchwork)
```

```{r}
source(here("scripts", "my_functions.R"))
```


## Datasets
```{r}
sparrows_raw <- read_tsv(here("data", "sparrows.txt"))
```

```{r}
sparrows <- sparrows_raw %>% 
  mutate(box_group = "Sparrows", 
         row_order = 1:nrow(.))
sparrows
```

#This code chunk adds two new columns, "box_group" and "row_order". 

### Figure 2: Outliers

```{r}
plot_box <- sparrows %>% 
  ggplot(aes(x = box_group, 
             y = wingcrd)) + 
  geom_boxplot(width = 0.25) + 
  labs( x = NULL, 
        y = "Wing length (mm)")
```

# This code chunk creates the boxplot seen in figure 2. 

```{r}
plot_cleveland <- sparrows %>% 
  ggplot(aes(x = wingcrd, 
             y = row_order)) + 
  geom_point() + 
  labs(x = "wing Length (mm)", 
       y = "Order of the data")
```

# This code chunk takes the above graph and instead plots the data on a cleveland plot to visualize the data points. This is to visualize the outlier points shown on the whisker graph above. Looking at the outliers there is only one that is really outside of the trend.  

```{r}
plot_box + plot_cleveland + coord_flip()
```

# This code chunk takes our previous two graphs and places them side by side. This is useful for comparing trends across multiple graphs. This is now used to see where the outliers sit between the rest of the data. Using a coord_flip() lets us line up the data for a better visual. 

#
***

### Figure 3: Outliers

```{r}
p_culmen <- sparrows %>% 
  ggplot(aes(y = row_order)) + 
  geom_point(aes(x = culmen)) + 
  theme(axis.title.y=element_blank(), 
        axis.text.y=element_blank(), 
        axis.ticks.y=element_blank())

p_nalospi <- sparrows %>% 
  ggplot(aes(y = row_order)) + 
  geom_point(aes(x = nalospi)) + 
  theme(axis.title.y=element_blank(), 
        axis.text.y=element_blank(), 
        axis.ticks.y=element_blank())

p_wt <- sparrows %>% 
  ggplot(aes(y = row_order)) + 
  geom_point(aes(x = wt)) + 
  theme(axis.title.y=element_blank(), 
        axis.text.y=element_blank(), 
        axis.ticks.y=element_blank())

p_wingcrd <- sparrows %>% 
  ggplot(aes(y = row_order)) + 
  geom_point(aes(x = wingcrd)) + 
  theme(axis.title.y=element_blank(), 
        axis.text.y=element_blank(), 
        axis.ticks.y=element_blank())

p_tarsus <- sparrows %>% 
  ggplot(aes(y = row_order)) + 
  geom_point(aes(x = tarsus)) + 
  theme(axis.title.y=element_blank(), 
        axis.text.y=element_blank(), 
        axis.ticks.y=element_blank())

p_head <- sparrows %>% 
  ggplot(aes(y = row_order)) + 
  geom_point(aes(x = head)) + 
  theme(axis.title.y=element_blank(), 
        axis.text.y=element_blank(), 
        axis.ticks.y=element_blank())

p_culmen + p_nalospi + p_wt + p_wingcrd + p_tarsus + p_head
```

# This code chunk recreates the Figure 3 seen in the study. The format for each of these graphs are very similar, as the only change that had to be made was on the x-variable. After the graphs are made they are then added together using the argument similar to the above. 

# The three plots with outlier data are: tarsus, head, and nalospi. Each of these have 1 or two points that are significantly outside the bounds of the other points. 

#
***

### Figure 5: distribution
```{r}
h1 <- sparrows %>% 
  filter(Month %in% 6:8) %>% 
                           ggplot() + 
                           geom_histogram(aes(x=wt), 
                                          binwidth = 0.5, 
                                          boundary = 0, 
                                          closed = "right", 
                                          color = "black") + 
  scale_x_continuous(breaks = seq(14, 28, by = 2))
h1
```

# This code chunk recreates the leftmost histogram in figure 5. The scale_x_continuous() argument labels the x-axis values by increments of 2. Leaving out this line instead labels it by 4. 

```{r}
h2 <- sparrows %>% 
  filter(Month %in% 6:8) %>%
  mutate(Month = case_when(
    Month == 6 ~ "June", 
    Month == 7 ~ "July", 
    Month == 8 ~ "August")) %>% 
  ggplot() + 
  geom_histogram(aes(x = wt), 
                 binwidth = 0.5, 
                 boundary = 0, 
                 color = "black") + 
  scale_x_continuous(breaks = seq(14, 28, by = 2)) + 
  facet_wrap(~ Month, ncol = 1)
h2
```

# This code chunk recreates the fright most histogram seen in figure 5 of the study. 

```{r}
h1 + h2
```

# This completes figure 5 by combining both sides of the histogram. 

#
***

# Figure 10: Relationships