---
title: "HW 14: Maps"
author: "Dylan Klingeman"
date: "`r format(Sys.time(), '%d %B %Y')`"
output: html_notebook
editor_options: 
  chunk_output_type: inline
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
library(here)
library(raster)
library(sf)
library(tidyverse)
library(ggthemes)
library(patchwork)
library(maptools)
library(ggmap)
library(marmap)
```

### Part 1: Outline maps

## US map

```{r}
usa <- map_data(map = "state") # loads data for us map

us_missouri <- map_data(map = "state", 
                        region = "missouri") # loads missouri data. 
us_other <- subset(x = usa, subset = region %in% c("alabama", 
                                                   "florida", 
                                                   "mississippi", 
                                                   "north carolina"))
```

```{r}
 ggplot() + 
  geom_polygon(data = us_other, 
               aes(x = long, 
                   y  = lat, 
                   group = group), 
               fill = "gray") + 
  geom_polygon(data = us_missouri, 
               aes(x = long, y = lat, group = group), 
               fill = "black") 
```

# this code chunk plots the longitude and latitude of 5 states onto a plot, creating a map. 

```{r}
us_map <-ggplot() + 
  geom_polygon(data = us_other, 
               aes(x = long, y = lat, group = group), 
               fill = "gray") + 
  geom_polygon(data = us_missouri, 
               aes(x = long, y = lat, 
                   group = group), 
               fill = "black") + 
  geom_polygon(data = usa, 
               aes(x = long,
                   y = lat, group = group), 
               fill = NA, 
               color = "black") + 
    theme_map() + 
  coord_fixed(1.3)

us_map
```

# creates a map of the 48 in-land us states. with Missouri colored black, and alabama, mississipi, north carolina, and florida colored grey. 


## Missouri Map
```{r}
missouri <- map_data("county", "missouri")
mo_scott <- subset(missouri, subregion %in% c("scott"))

mo_map <- ggplot() + 
  geom_polygon(data = missouri, aes(x = long, y = lat, group = group), fill = NA, color = "black") + 
  geom_polygon(data = mo_scott, aes(x = long, y = lat, group = group), fill = "black") + 
  theme_map() + 
  coord_fixed(1.3)

mo_map
```

# Creates a map of missouri's counties and highlights scott county in black. 

```{r}
us_map + mo_map + plot_layout(ncol = 2, widths = c(1.5, 1))
```

# pastes the above two maps next to each other side by side (2 columns). The width of the first map is increased to 1.5x size. 

## Outline maps: your turn

```{r}
missouri <- map_data("county", "missouri")
mo_daviess <- subset(missouri, subregion %in% c("daviess"))
mo_knox <- subset(missouri, subregion %in% c("knox"))
mo_cape_girardeau <- subset(missouri, subregion %in% c("cape girardeau"))

mo_map2 <- ggplot() + 
  geom_polygon(data = missouri, aes(x = long, y = lat, group = group), fill = NA, color = "black") + 
  geom_polygon(data = mo_cape_girardeau, aes(x = long, y = lat, group = group), fill = "black") + 
  geom_polygon(data = mo_daviess, aes(x = long, y = lat, group = group), fill = "gray70") + 
  geom_polygon(data = mo_knox, aes(x = long, y = lat, group = group), fill = "gray70") + 
  coord_fixed(1.3) + 
  theme_map()

mo_map2
```

# This code chunk makes a missouri map similar to the above, but highlights different counties. For the 2 name counties: Daviess and Knox. As wella s the birth county, cape girardeau. 

```{r}
us_map2 <-ggplot() + 
  geom_polygon(data = us_missouri, 
               aes(x = long, y = lat, 
                   group = group), 
               fill = "black") + 
  geom_polygon(data = usa, 
               aes(x = long,
                   y = lat, group = group), 
               fill = NA, 
               color = "black") + 
    theme_map() + 
  coord_fixed(1.3)

us_map2
```

# Creates the us map with only my birth state highlighted, missouri. 

```{r}
mo_map2 + us_map2 + plot_layout(ncol = 2, widths = c(1, 1.5))
```

# The two maps froma bove attached together. 

#

### Part 2: Choropleth maps


```{r}
library(sf)
library(tidyverse)
```

```{r}
skin_cancer_raw <- 
  read_delim(file = here::here("data", "skin_cancer.txt"), 
             delim = " ", 
             skip = 7,
             trim_ws = TRUE)
```
# loads in the skin_cancer.txt data

```{r}
skin_cancer <- skin_cancer_raw %>% 
  filter(State != "Wash,D.C.") %>% 
  mutate(State = case_when(
    State == "NorthCarolina" ~ "North Carolina", 
    State == "SouthCarolina" ~ "South Carolina", 
    State == "NorthDakota" ~ "North Dakota", 
    State == "SouthDakota" ~ "South Dakota", 
    State == "NewHampshire" ~ "New Hampshire", 
    State == "NewJersey" ~ "New Jersey",
    State == "NewMexico" ~ "New Mexico",
    State == "MewYork" ~ "New York", 
    State == "RhodeIsland" ~ "Rhode Island",
    State == "WestVirginia" ~ "West Virginia",
    TRUE ~ State
  ))
```

# Relabels the states, removing spaces, etc. 

```{r}
mean_mort <-  mean(skin_cancer$Mort)

skin_cancer_adj <- skin_cancer %>% 
  mutate(relative_mort = Mort - mean_mort)
```

# calculates the mean mortality of each state. 

```{r}
lower_48 <- c("Alabama","Arizona","Arkansas","California","Colorado","Connecticut","Delaware","Florida","Georgia","Idaho","Illinois","Indiana","Iowa","Kansas","Kentucky","Louisiana","Maine","Maryland","Massachusetts","Michigan","Minnesota","Mississippi","Missouri","Montana","Nebraska","Nevada","New Hampshire","New Jersey","New Mexico","New York","North Carolina","North Dakota","Ohio","Oklahoma","Oregon","Pennsylvania","Rhode Island","South Carolina","South Dakota","Tennessee","Texas","Utah","Vermont","Virginia","Washington","West Virginia","Wisconsin","Wyoming")
```

# Removes Hawaii, D.C, Alaska, and us territories. 

```{r}
states <- st_read(here::here("data/shape_files", "cb_2017_us_state_500k.shp"), 
                  stringsAsFactors = TRUE)
```

```{r}
states_df <- states %>% 
dplyr::filter(NAME %in% lower_48)
```

```{r}
str(states_df$STATEFP)
```

```{r}
state_levels <- levels(states_df$STATEFP)
name_levels <- levels(states_df$NAME)
```

```{r}
library(tidyverse)
fips <- read_csv(here::here("data", "state_fips.csv")) %>% 
  select(-abbr) %>% 
  rename(STATEFP = FIPS) %>% 
    mutate(STATEFP = factor(STATEFP, levels = state_levels))

```

# This was messing up, took awhile to figure it out. The problem was state_levels vector wasn't named properly! 

# This combines and renames some columsn so that the data can be combined into one tibble. 

```{r}
skin_cancer_adj <- left_join(skin_cancer_adj, fips)

states_df <- left_join(states_df, skin_cancer_adj)
```

```{r}
ggplot(states_df) + geom_sf(aes(fill = relative_mort)) + 
  scale_fill_viridis_c(name = "Skin Cancer, \nRelative Mortality", 
                       option = "inferno") + 
  coord_sf(crs = st_crs(5070)) + 
  theme_bw() + 
  theme(legend.position = "bottom")
```

# Makes a map graph with a gradient to skin cancer mortality by state. The color scheme is set to the "inferno" viridis setting. 

#

## Choropleth map: your turn. 

```{r}
raw_lyme_disease<- read_csv(here::here("data", "lyme_disease.csv"), 
                            skip = 1, 
                            na = c("-"))
raw_lyme_disease
```

```{r}
wrangled_lyme_disease <- raw_lyme_disease %>% 
 mutate(`2017` = `2017_Confirmed` + `2017_Probable`) %>% 
  select(State: "2016", "2017") %>% 
  pivot_longer("2007":"2017", names_to = "year", values_to = "cases")%>% 
  group_by(State) %>% 
  summarise(mean_cases = mean(cases), 
            log_mean = log10(mean_cases), 
            .groups = "drop")
```

# wrangles the lyme disease dataset, collapsing some columns and combining others .

```{r}
lyme_disease_adj <- left_join(wrangled_lyme_disease, fips)

lyme_disease_df <- left_join(states_df, wrangled_lyme_disease)
```
# leftjoins the datasets together for graphing. 

```{r}
ggplot(lyme_disease_df) + 
  geom_sf(aes(fill = log_mean)) + 
  scale_fill_viridis_c(name = "Lyme Disease, \nCases in State", 
                       option = "cividis", labels = c("<1", "10", "100", "1000", "5000")) +
  coord_sf(crs = st_crs(5070)) + 
  theme_bw() + 
  theme(legend.position = "bottom")
```


#

## Part 3: Dot distribution maps

```{r}
library(raster)
library(tidyverse)
library(ggthemes)
library(ggmap)
```

```{r}
min_long <- -126
max_long <- -104
min_lat <-  31
max_lat <- 50

set.seed(9618973)
```

# sets the seed so all fhte values are consistent, and then sets guidelines for some variables. 

```{r}
bigfoot <- read_csv(here::here("data", "bigfoot.csv"))

bigfoot <- bigfoot %>% 
  filter(long >=min_long & long <= max_long, 
         lat >= min_lat & lat <= max_lat) %>% 
  sample_n(300) %>% 
  mutate(name = "Gigantopithecus") %>% 
  dplyr::select(name, long, lat)
```

```{r}
bears <- read_csv(here::here("data", "bears.csv"))

both_species <- bind_rows(bigfoot, bears)

head(both_species)
```

# loads in the bears dataset, then binds the bigfoot and bears together. Compares the first couple rows of data. 

```{r}
tail(both_species)
```

# compares the last couple rows of data between the two species. 

```{r}
base = get_stamenmap(bbox = c(min_long, 
                              min_lat, 
                              max_long, 
                              max_lat), 
                     zoom = 4, 
                     maptype = "terrain-background")

ggmap(base)
```

# Plots a map of the western US. showing forest ranges and land with different color schemes. 

```{r}
base <- ggmap(base)

base + geom_point(data = bigfoot, 
                  aes(x = long, 
                      y = lat))
```

# plots the 300 bigfoot data points from the earlier seed. 

```{r}
base + geom_point(data = bears, 
                  aes(x = long, 
                      y = lat))
```

# plots the black bear data points, to show as a comparison tot he bigfoot data. 

```{r}
base + geom_point(data = both_species,
                  aes(x = long,
                      y = lat,
                      shape = name,
                      fill = name),
                  size = 2) +
  scale_shape_manual(values = c(22:23)) +
  labs(x = "Longitude",
       y = "Latitude",
       fill = "Species",
       shape = "Species")
```

# Combines the two datasets and assigns themr espective colors to a legend. 

## Distribution map: your turn

