---
title: "HW 14: Maps"
author: "Dylan Klingeman"
date: "`r format(Sys.time(), '%d %B %Y')`"
output: html_notebook
editor_options: 
  chunk_output_type: inline
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
library(here)
library(raster)
library(sf)
library(tidyverse)
library(ggthemes)
library(patchwork)
library(maptools)
library(ggmap)
library(marmap)
```

### Part 1: Outline maps

## US map

```{r}
usa <- map_data(map = "state") # loads data for us map

us_missouri <- map_data(map = "state", 
                        region = "missouri") # loads missouri data. 
us_other <- subset(x = usa, subset = region %in% c("alabama", 
                                                   "florida", 
                                                   "mississippi", 
                                                   "north carolina"))
```

```{r}
 ggplot() + 
  geom_polygon(data = us_other, 
               aes(x = long, 
                   y  = lat, 
                   group = group), 
               fill = "gray") + 
  geom_polygon(data = us_missouri, 
               aes(x = long, y = lat, group = group), 
               fill = "black") 
```

# this code chunk plots the longitude and latitude of 5 states onto a plot, creating a map. 

```{r}
us_map <-ggplot() + 
  geom_polygon(data = us_other, 
               aes(x = long, y = lat, group = group), 
               fill = "gray") + 
  geom_polygon(data = us_missouri, 
               aes(x = long, y = lat, 
                   group = group), 
               fill = "black") + 
  geom_polygon(data = usa, 
               aes(x = long,
                   y = lat, group = group), 
               fill = NA, 
               color = "black") + 
    theme_map() + 
  coord_fixed(1.3)

us_map
```

# creates a map of the 48 in-land us states. with Missouri colored black, and alabama, mississipi, north carolina, and florida colored grey. 


## Missouri Map
```{r}
missouri <- map_data("county", "missouri")
mo_scott <- subset(missouri, subregion %in% c("scott"))

mo_map <- ggplot() + 
  geom_polygon(data = missouri, aes(x = long, y = lat, group = group), fill = NA, color = "black") + 
  geom_polygon(data = mo_scott, aes(x = long, y = lat, group = group), fill = "black") + 
  theme_map() + 
  coord_fixed(1.3)

mo_map
```

# Creates a map of missouri's counties and highlights scott county in black. 

```{r}
us_map + mo_map + plot_layout(ncol = 2, widths = c(1.5, 1))
```

# pastes the above two maps next to each other side by side (2 columns). The width of the first map is increased to 1.5x size. 