---
title: "HW06 Part 1: Complete the sections"
author: "Dylan Klingeman"
date: "`r format(Sys.time(), '%d %B %Y')`"
output: html_notebook
editor_options: 
  chunk_output_type: inline
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

- Change "your name" in the YAML header above to your name.

- As usual, enter the examples in code chunks and run them, unless told otherwise.

## Chapter 10: Tibbles

Read [R4ds Chapter 10: Tibbles](https://r4ds.had.co.nz/tibbles.html), sections 1-3.

### 10.1: Introduction

Load the tidyverse package. 


```{r}
library(tidyverse)
```

Loads the tidyverse package, to be used in later sections. 

#

### 10.2: Creating tibbles

Enter your code chunks for Section 10.2 here.

Describe what each chunk code does. 


```{r}
as_tibble(iris)
```

Creates a tible for the iris dataset. A tible is similar to a data frame, but with some notable differences in it's execution. 

```{r}
tibble(
  x = 1:5, 
  y = 1, 
  z = x ^ 2 + y
)
```

Creates a simple table with 5 rows, showing some math for example. 

```{r}
tb <- tibble(
  ':)' = "smile", 
  ' ' = "space", 
  '2000' = "number"
)
tb
```

Creates "tb" which is a table with some smalls ample information, within a tibble. 

```{r}
tribble( 
  ~x, ~y, ~z, 
  #--/--/----
  "a", 2, 3.6, 
  "b", 1, 8.5
  )
```
Creates a table and assigns values to an x, y, and z axis. 

#

### 10.3: Tibbles vs data.frame

Enter your code chunks for Section 10.2 here.

Describe what each chunk code does. 


```{r}
df <- tibble(
  x = runif(5), 
  y = rnorm(5)
)
df$x

df[["x"]]

df[[1]]
```


This step creates a tibble and then shows how to pull out specific pieces within it. In this example it shows how to get differnet places in the x variable. 

#

### 10.4: Not required

#### Section 10.5 Questions

Answer the questions *completely.* Use code chunks, text, or both, as necessary.

**1:** How can you tell if an object is a tibble? (Hint: try printing `mtcars`, which is a regular data frame). Identify at least two ways to tell if an object is a tibble. *Hint:* What does `as_tibble()` do? What does `class()` do? What does `str()` do?


The easiest way to tell if an object is a tibble is looking at the number of rows that appear when looking at the data. A tibble will only show the 1st 10 rows out of the data set, but a normal data set will show all of the rows. Another feature of tibble is that it doesnt change: input type, names of variables, or create names for rows. 



**2:** Compare and contrast the following operations on a data.frame and equivalent tibble. What is different? Why might the default data frame behaviours cause you frustration?

```{r}
df <- data.frame(abc = 1, xyz = "a")
df$x
df[, "xyz"]
df[, c("abc", "xyz")]
```


```{r}
dt <- tibble( 
  "abc" = 1, 
  "xyz" = "a"
  )
dt$xyz
dt[["xyz"]]
dt[[2]]
```

# Creates a tibble for the above data frame and saves it to a variable "dt". Then displays the 2nd variable in 3 differant ways for example. 

Between the two, when searching for a specific variable it is much easier to do on a tibble. For a data frame you have to state the specific row or column that the variable is in, then name it or state the x/y axis point. With a tibble it is much easier to state the variable that you want at that time with $. The data frame is still able to acquire this information, but is more complex to have the same result. 

## Chapter 11: Importing data

Read [R4ds Chapter 11: Data Import](https://r4ds.had.co.nz/data-import.html), sections 1, 2, and 5.

### 11.1 Introduction

Nothing to do here unless you took a break and need to reload `tidyverse`.

```{r}
library(tidyverse)
```

### 11.2 Getting started.

Do *not* run the first code chunk of this section, which begins with `heights <- read_csv("data/heights.csv")`. You do not have that data file so the code will not run.

Enter and run the remaining chunks in this section.

```{r}
read_csv("a, b, c
         1, 2, 3
         4, 5, 6")
```

# An example tibble that shows how the read_csv function works. Uses the first line to title the column names with each "," separating the names. 

```{r}
read_csv("The first line of metadata
         The second line of metadata
         x, y, z
         1, 2, 3", skip=2)
```

# An example tibble that shows off the "skip = n" function. This function will skip "n" lines before it reads the line to pick a title. In this case it skips the two lines of text and uses "x, y, z" as the names. 

```{r}
read_csv("1,2,3\n4,5,6", col_names = FALSE)
```

# This example creates a tibble, but prevents the read_csv from takign the column names. Instead they are given "Xn" column names, and the first line iwll instead be listed as a row. 

```{r}
read_csv("1,2,3\n4,5, 6", col_names = c("x", "y", "z"))
```

# This example shows how to use col_names within a tibble to list the column name, and have the first line become the first row. 

#### 11.2 Questions

**1:** What function would you use to read a file where fields were separated with "|"?

To read a file with pipes we sould use the "read_delim(). 



**2:** (This question is modified from the text.) Finish the two lines of `read_delim` code so that the first one would read a comma-separated file and the second would read a tab-separated file. You only need to worry about the delimiter. Do not worry about other arguments. Replace the dots in each line with the rest of your code. 

# Comma-separated
`file <- read_delim("file.csv", delim = ",")'`

# Tab-separated
`file <- read_delim("file.csv", "   ")`



**3:** What are the two most important arguments to `read_fwf()`? Why?

file.csv: Provides data for read_fwf to extract, needed for any function. 
fwf_width/fwf_position: Used to specifiy points and positions in the table to work with. 



**4:** Skip this question


**5: ** Identify what is wrong with each of the following inline CSV files. What happens when you run the code?

```{r}
read_csv("a,b\n1,2,3\n4,5,6")
read_csv("a,b,c\n1,2\n1,2,3,4")
read_csv("a,b\n\"1")
read_csv("a,b\n1,2\na,b")
read_csv("a;b\n1;3")
```

Line 1: A column is missing as well as two data points: 3 and 6. This happens because there is no name in the first line. 
  Potential fix:  read_csv("a,b, c\n1, 2, 3\n4, 5, 6")
  
Line2: There is a missing input on line 2 as well as an extra input on line 4. 
  Potential fix 1: read_csv("a,b,c\n1,2, 3\n1,2,3")
  Potential fix 2: read_csv("a,b,c,d\n1,2\n1,2,3,4")
  
Line3: There are two columns, but only 1 point of data. This gives a null on the 2nd row 2nd position. Either adding a point to this position or removing a column would fix the issue. There is also separated backticks on the lines. 
  Potential fix1: read_csv("a\n\"1")
  Potential fix2: read_csv("a, b\n1, 2")
  
Line4: This one is strange as it lists letters in teh data frame. These are usually used to refer to points/positions and placed inside of "". Having these here seems completely unintuitive. 
  Potential fix1: read_csv("a,b\n1,2")
  
Line5: The data is being separated by semicolons instead of commas, which causes the data to not separate properly. Either changing read_csv to read_csv2 or changing the semicolons out for commas would fix the issue. 
  Potential fix 1: read_csv2("a;b\n1;3")
  Potential fix 2: read_csv("a, b\n1,3")

#

### 11.3 and 11.4: Not required

### 11.5: Writing to a file

Just read this section. You may find it helpful in the future to save a data file to your hard drive. It is basically the same format as reading a file, except that you must specify the data object to save, in addition to the path and file name.

### 11.6 Not required

## Chapter 18: Pipes

Read [R4ds Chapter 18: Pipes](https://r4ds.had.co.nz/pipes.html), sections 1-3.

Nothing to do otherwise for this chapter. Is this easy or what?

**Note:** Trying using pipes for all of the remaining examples. That will help you understand them.

## Chapter 12: Tidy Data

Read [R4ds Chapter 12: Tidy Data](https://r4ds.had.co.nz/tidy-data.html), sections 1-3, 7. 

### 12.1 Introduction

Nothing to do here unless you took a break and need to reload the `tidyverse.`

### 12.2 Tidy data

Study Figure 12.1 and relate the diagram to the three rules listed just above them. Relate that back to the example I gave you in the notes. Bear this in mind as you make data tidy in the second part of this assignment.

You do not have to run any of the examples in this section.

### 12.3

Read and run the examples through section 12.3.1 (gathering), including the example with `left_join()`. We'll cover joins later.

```{r}
table4a
```

example table that shows a layout of a tibble. 

```{r}
table4a %>% 
  pivot_longer(c('1999', '2000'), names_to = "year", values_to = "cases")
```

modifies the table above by pivoting the data around. In this case it takes the 2 columns above and merges them while applying the name "year" to them. In addition it moves the values given into their own columns and assigns the name "cases". 

```{r}
table4b %>% 
  pivot_longer(c('1999', '2000'), names_to = "year", values_to = "population")
```

Similar to the example above, this modifies the table4b and arranges the data so that the value is separated from it's description (years to population in this case).

```{r}
tidy4a <- table4a %>% 
  pivot_longer(c('1999', '2000'), names_to = "year", values_to = "cases")
tidy4b <- table4b %>% 
  pivot_longer(c('1999', '2000'), names_to = "year", values_to = "population")
left_join(tidy4a, tidy4b)
```

This example does everything all in one! It repeats the modifications in the two examples above, but also introduces the joining argument, which combines data sets together. By doing "left_join" it attaches the two vectors together and makes a combined table. 

```{r}
table2
```

An example table from the data base, showing TB cases around the world. 

```{r}
table2 %>% 
  pivot_wider(names_from = type, values_from = count)
```

This example introduces us to "pivot_wider" which is an argument like pivot_longer from above. Pivot_wider works to combine data within a column, and use it to create new columns to reduce the length of the data set. In this example it cuts the armount of rows necessary in half. 

#### 12.3 Questions

**2:** Why does this code fail? Fix it so it works.

```{r}
table4a %>% 
  gather(1999, 2000, key = "year", value = "cases")
#> Error in inds_combine(.vars, ind_list): Position must be between 0 and n
```

When comparing it to the code from above, we can see several changes that cause this code to not work. The numvers 1999 and 2000 need to be within "" to be recognized as variables. In addition the pivot_longer has been changed to "gather", which will not work in this context. The other change is names_to into "key" which should be reverted. Doing these steps will give a functional tibble as shown below. 

```{r}
table4a %>% 
  gather("1999, 2000", key = "year", value = "cases")
```

```{r}
table4a %>% 
  pivot_longer(c("1999", "2000"), names_to = "year", values_to = "cases")
```

Example fix to problem above. 

That is all for Chapter 12. On to the last chapter.


## Chapter 5: Data transformation

Read [R4ds Chapter 5: Data Transformation](https://r4ds.had.co.nz/transform.html), sections 1-4.

Time to [get small.](https://www.youtube.com/watch?v=GOrdzCHnpw4) 

### 5.1: Introduction

Load the necessary libraries. As usual, type the examples into and run the code chunks.


```{r}
library(nycflights13)
library(tidyverse)
```

### 5.2: Filter rows with `filter()`

Study Figure 5.1 carefully. Once you learn the `&`, `|`, and `!` logic, you will find them to be very powerful tools.


#### 5.2 Questions

**1.1:** Find all flights with a delay of 2 hours or more.

```{r}
filter(flights, !(arr_delay <= 120 | dep_delay <= 120))
```



This creates a filter that will show only data within "flights" where the arrival or departure delay were greater than 2 hours (120 minutes). The | creates the "or" and the ! says not, which then removes the data it is then told to.  

**1.2:** Flew to Houston (IAH or HOU)

```{r}
filter(flights, dest == "IAH" | dest == "HOU")
```

Filters for flights that had a destination of Houston which is "IAH" or "HOU". Backticks are key to looking up word variables. 

**1.3:** Were operated by United (UA), American (AA), or Delta (DL).

```{r}
filter(flights, carrier %in% c("UA", "AA", "DL"))
```

Filters for flights that had the above carriers: UA, AA, or DL. Again note the quotation around the names given. Also using the "%in%" argument to search for specific variables. in this case we use it on the three names above. 

**1.4:** Departed in summer (July, August, and September).

```{r}
filter(flights, month %in% c(7:9))
```

Filters for flights in July, August and September (7th, 8th, and 9th month of theyear respectively). This is done again usign the %in% and a colon. Originally I thought to list out the numbers, 7, 8, 9, but using the colon saves coding space. 

**1.5:** Arrived more than two hours late, but didn’t leave late.


```{r}
filter(flights, arr_delay >= 120 & dep_delay <= 0)
```

Filters for flights that had no departure delay, or left early that arrived 2 hours or more late. I had some difficulty getting to work this in a format with the ! like above. 

**1.6:** Were delayed by at least an hour, but made up over 30 minutes in flight. This is a tricky one. Do your best.


```{r}
filter(flights, dep_delay >=60 & (arr_delay - dep_delay) < -30)
```

Filters for flights that were delayed by at least an hour, as well as made up at least 30 minutes ruing the flight. This is done by subtracting the arrival delay from the departure delay, then filtering by differences of 30 or higher. I initially tried to just show arrival delays that were less than 30 minutes, but this was not able to pick up the flights that had very large initial delays. 

**1.7:** Departed between midnight and 6am (inclusive)


```{r}
filter(flights, dep_time %in% (0:600))
```

Filters for flights that departed between midnight (000) and 6 am (600). I initially did this b doing a greater than statement, with an & as well as a lesser than. However, doing it like this was much shorter and neater to do the same thing. 


**2:** Another useful dplyr filtering helper is `between()`. What does it do? Can you use it to simplify the code needed to answer the previous challenges?

"Between()" is an argument that filters data that is only between two variables. This can be used similarly to the %in% questions above, replacing that argument with a colon. Using the between argument would not be inclusive, but it should otherwise be similar. 

1.4: 
```{r}
filter(flights, between(month, 7, 9))
```

1.7: 
```{r}
filter(flights, between (dep_time, 0, 600))
```


**3:** How many flights have a missing dep_time? What other variables are missing? What might these rows represent?

```{r}
filter(flights, is.na(dep_time))
```

There are 8,225 flights that are missing their departure times, as well as their: departure delay, arrival times, arrival delay, and air times every single time. The tail number of the plane was also missing sometimes, but rarely. These are most likely flights that were canceled. 


**4:** Why is `NA ^ 0` not missing? Why is `NA | TRUE` not missing? Why is `FALSE & NA` not missing? Can you figure out the general rule? (`NA * 0` is a tricky counterexample!)


```{r}
NA^0
```

This will always produce 1, because anyting to the power of 0 will produce 1. 

```{r}
NA|TRUE
```

This returns "TRUE" because it is assignign true to the NA variable. This would do the same thing with anything placed there as below. 

```{r}
3|TRUE
```
 
 
```{r}
FALSE & NA
```

This is similar to the true statement, by assigning the variable to be false it will return NA no matter what you put in there as shown below. 

```{r}
FALSE & 3
```

The general rule that I take from this is that, assigning a condition to a variable works on most anything. If needed this can be used to mark something in absolute. Another take from this is that NA is more of an answer to an output than it is a variable. Manipulating it with conditions is easy as shown above, but this is taking it out of context of it's purpose. 

**Note:** For some context, see [this thread](https://blog.revolutionanalytics.com/2016/07/understanding-na-in-r.html)


### 5.3 Arrange with `arrange()`

```{r}
arrange(flights, year, month, day)
```

This arranges the order of flights so that they are displayed in order of their year, month, and days. 

```{r}
arrange(flights, desc(dep_delay))
```

This arranges the flights in order of decreasing departure delay. 

```{r}
df <- tibble(x = c(5, 2, NA))
arrange(df, x)
```

This creates a tibble with 2 numbers out of order. Then the arrange argument is used to enter the numbers in a correct order. 

```{r}
arrange(df, desc(x))
```

This takes the above tibble  and lists the numbers in descending order. 

#### 5.3 Questions

**1:** How could you use `arrange()` to sort all missing values to the start? (Hint: use is.na()). **Note:** This one should still have the earliest departure dates after the `NA`s. *Hint:* What does `desc()` do?

```{r}
arrange(flights, desc(year, month, day,  is.na(year, month,day, dept_time)))
```


    

**2:** Sort flights to find the most delayed flights. Find the flights that left earliest. 

This question is asking for the flights that were most delayed (left latest after scheduled departure time) and least delayed (left ahead of scheduled time).

```{r}
arrange(flights, desc(dep_delay))
```

Arranges flights in order of how long they were delayed, with the most at the top. 1301 minutes is over 20 hours!

```{r}
arrange(flights, -desc(dep_delay))
```

Arranges flights in order of how long they were delayed, with the least at the top. 43 minutes ahead of schedule is quite alot, I imagine quite a few missed that flight. 

**3:** Sort flights to find the fastest flights. Interpret fastest to mean shortest time in the air.

```{r}
arrange(flights, -desc(air_time))
```

Arranges flights in order of how long they took to land. The fastest was 20 minutes. 

*Optional challenge:* fastest flight could refer to fastest air speed. Speed is measured in miles per hour but time is minutes. Arrange the data by fastest air speed.

```{r}
arrange(flights, desc(distance/air_time))
```

This arranges flights in order of the highest distance traveled per air_time, which translates to the fastest speed. The top row traveled 762 miles in 65 minutes. Dividing this out and then multiplying by sixty gives over 700 miles per hour! This seems extremely fast. 

**4:** Which flights travelled the longest? Which travelled the shortest?

```{r}
arrange(flights, desc(distance))
```

This arranges flights in descending order of how many miles they traveled. 

```{r}
arrange(flights, -desc(distance))
```

This arranges flights in ascending order of how many miles they traveled. 

### 5.4 Select columns with `select()`

#### 5.4 Questions

**1:** Brainstorm as many ways as possible to select `dep_time`, `dep_delay`, `arr_time`, and `arr_delay` from flights. Find at least three ways.


**2:** What happens if you include the name of a variable multiple times in a `select()` call?



**3:** What does the `one_of()` function do? Why might it be helpful in conjunction with this vector?

`vars <- c("year", "month", "day", "dep_delay", "arr_delay")`



**4:** Does the result of running the following code surprise you? How do the select helpers deal with case by default? How can you change that default?

`select(flights, contains("TIME"))`


