---
title: "HW 13: Strings and things"
author: "Dylan Klingeman"
date: "`r format(Sys.time(), '%d %B %Y')`"
output: html_notebook
editor_options: 
  chunk_output_type: inline
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
library(tidyverse)
library(here)
library(ggthemes)
library(gghighlight)
```

### Part 1: DNA strings

```{r}
gliadin_data <- read_lines(here("data", "beta_gliadin.txt"), skip = 1)
gliadin_data
```

# Imports the beta gliadin dataset and removes the first line. 

```{r}
gliadin_wrangled <- str_replace_all(gliadin_data, c("\\d" = "", "\\s" = "")) %>%  # removes all digits and spaces
  
str_c(collapse = "") %>% # collapses into a single string

str_to_upper() # converts to uppercase letters

str_length(gliadin_wrangled) # measures string length

gliadin_wrangled
```

# Put all the steps into one block to save space. Will try to do this as much as possible going forward. 

```{r}
str_count(gliadin_wrangled, "[A.C.T.G]")   # All are A, C, T, G
str_count(gliadin_wrangled, "ATG") # 66 matches
str_count(gliadin_wrangled, "TGA|TAA") # 99 matches
```

# Counts totals, had to use only one set of quotations to make the final line work. 

```{r}
caat_box <- "CCAAT"
tata_box <- "CTATAA"

str_count(gliadin_wrangled, caat_box) # 4 matches
str_count(gliadin_wrangled, tata_box) # 1 match

caat_seq <- "CCAAT.{100,150}ATG" # like in 14.3.4
str_count(gliadin_wrangled, caat_seq) # 2 matches

tata_seq <- "CTATAA.{50,100}ATG" # 1 match
str_count(gliadin_wrangled, tata_seq)
```

# tata seq was returning 0 matches, fixed typo from "CATATAA" -> "CTATAA"

```{r}
caat_seqs_all <- str_extract_all(gliadin_wrangled, caat_seq, simplify = TRUE) # saves all caat sequences

caat_seqs_all
str_length(caat_seqs_all) # sequence 1 is 139, sequence 2 is 148. 

str_which(caat_seqs_all, tata_seq) # sequence1 contains the tata sequence. 

promotor <- str_subset(caat_seqs_all, tata_seq) # prints the entire string 1.
promotor
```

# 

### Part 2: Ants

```{r}
ant_data <- read_csv(here("data", "ants.csv"))
head(ant_data)
```

# imports the data and shows first couple lines. 

```{r}
ant_wrangled <- ant_data %>%
   pivot_longer(names_to = "species", # Moves names to species
                values_to = "abundance", # Move value to abundance column
                matches("[A-Z].+\\.[a-z]+", # Checks that names are in uppercase then lowercase format. 
                        ignore.case = FALSE)) %>% # Case sensitive. 
  select(year, species, abundance) %>% 
   mutate(species = str_replace(species,
                                "(.{4}).*\\.(.{4}).*",
                                "\\1\\2"))  %>% 
  mutate(year = factor(year, ordered = TRUE, levels = c(2010, 2011))) %>% 
  group_by(year, species) %>% 
  summarise(total_abundance = sum(abundance))
ant_wrangled

```

# had to mess around with the "." after the "a-z" to make it work

# "(.{4}).*\\.(.{4}).*" Captures the first 4 letters of the first word, spaces to the ext word, and then captures 4 letters. 

# the two parts separated by the \\ can be back referenced with 1 and 2 respectivelyh. 

## Was stuck until i removed ang_wrangled from the summarise argument. Was adding all of the totals together and making very large numbers. 

```{r}
ggplot(ant_wrangled, aes(x = reorder(species, total_abundance), y = total_abundance), group = species) + 
  geom_line() + 
  geom_point(aes(shape = year), size = 2) + 
  coord_flip() + 
  labs(x = NULL, y = "Total Abundance", shape = "Year") + 
  theme_minimal()
  
```

#

### Part 3: Featuring Phlox flowers. 

# First data set

```{r}
flower_data <- read_csv(here("data", "flower_size.csv"))
flower_data
```

```{r}
wrangled_flower_data <- flower_data %>% 
  filter(!is.na("Flower number")) %>% # keeps na values
  select(species = "Species",cor_length =  'Corolla length (cm)',throat_width =  `Corolla width throat (cm)`) %>% # renames columns
separate(species, into = c("genus", "species"), extra = "drop") %>% 
  group_by(genus, species) %>% 
  summarise(mean_length = mean(cor_length, na.rm = TRUE), # mean length
            mean_width = mean(throat_width, na.rm = TRUE), # mean width
            max_length = max(cor_length, na.rm = TRUE), # max length
            max_width = max(throat_width, na.rm = TRUE) # max width
            , .groups = 'drop') # fixes group argument

wrangled_flower_data
```

# Second data set

```{r}
pollinators_raw <- read_csv(here("data", "flower_pollinators.csv"))
```

```{r}
pollinators <- pollinators_raw %>% 
   select(-c(starts_with("Source"), Color)) %>%
  filter(!is.na(Pollinator)) %>% 
  separate(Species, into = c("genus", "species"), extra = "drop") %>% 
  separate(Pollinator, into = "Syndrome", sep = " ")
  
pollinators
```

```{r}
phlox <- inner_join(wrangled_flower_data, pollinators)
phlox
```

```{r}
ggplot(data = phlox) + 
  geom_point(aes( x = mean_length, y = mean_width, color = Syndrome)) 
```

```{r}
phlox_mutated <- phlox %>% 
  mutate(Syndrome = str_replace(Syndrome, "(^be.*e$)", "\\1s"),
   Syndrome = str_replace(Syndrome, ".+l.?es", "other insects"))
```
# groups bee/beetle into bees/beetles respectively. 

# Removes Beetles and flies, but replaces them with other insects. 

```{r}
ggplot(data = phlox_mutated) + 
  geom_point(aes( x = mean_length, y = mean_width, color = Syndrome)) + 
  labs(x = "Mean length", y = "Mean width")
```

# graph with bees/beetles. 

```{r}
phlox %>% 
  group_by(Syndrome) %>% 
  count(Syndrome)
```


```{r}
phlox_other <- phlox_mutated %>% 
  mutate(Syndrome = str_replace(Syndrome,".+1.?es", "other insects"))
phlox_other %>% 
  ggplot() + 
  geom_point(aes(x = mean_length, y = mean_width, color = Syndrome)) + 
  labs(x = "Mean length", y = "Mean width")
```

# Graph With other insects from above. 

```{r}
(phlox_categorised <- phlox %>%
  mutate(Syndrome = str_replace(Syndrome, "(^be.*e$)", "\\1s"),
         Syndrome = str_replace(Syndrome, ".+l.?es", "other insects"),
         syndrome_group = case_when(
           Syndrome == "autogamous" ~ "Autogamous",
          Syndrome == "bat" ~ "Mammal",
          Syndrome == "hummingbird" ~ "Bird",
          TRUE ~ "Insect"
         )))

ggplot(data = phlox_syndrome) + 
  geom_point(aes(x = mean_length, y = mean_width, color = syndrome_group)) + 
  labs(x = "Mean length", y = "Mean width", color = "Syndrome group") + 
  gghighlight(syndrome_group == "Mammal" | syndrome_group == "Bird", use_direct_label = FALSE)
```

# Cuts down to 4 groups on the legend. 

### Part 4: Customize your plots

# Plot 1

```{r}
phlox_categorised %>% 
  ggplot() + 
  geom_point(aes(x = max_length, y = max_width, shape = syndrome_group, color = syndrome_group)) + 
  labs( x = "Max length", y = "Max width", color = "Syndrome", shape = "Syndrome") + 
  scale_color_brewer(type = "qual", palette = "Accent")
```

# Plot 2 

```{r}
phlox_categorised %>% 
  ggplot() + 
  geom_point(aes(x = max_length, y = max_width, shape = syndrome_group, color = syndrome_group), size = 1.5) + 
  labs( x = "Max length", y = "Max width", color = "Syndrome", shape = "Syndrome") + 
  scale_color_brewer(type = "Diverging", palette = "BrBG")
```

# Plot 3 

```{r}
phlox_categorised %>% 
  ggplot() + 
  geom_point(aes(x = max_length, y = max_width, shape = syndrome_group, fill = syndrome_group), size = 2) + 
  labs( x = "Max length", y = "Max width", shape = "Syndrome", fill = "Syndrome") + 
  scale_fill_viridis_d(option = "cividis") + 
    scale_shape_manual(values = c(22:25)) +
  theme_classic()
```

# Plot 4

```{r}
phlox_categorised %>% 
  ggplot() + 
  geom_point(aes(x = max_length, y = max_width, shape = syndrome_group, fill = syndrome_group), size = 2) + 
  labs( x = "Max length", y = "Max width", shape = "Syndrome", fill = "Syndrome") + 
  scale_fill_viridis_d(option = "viridis") + 
    scale_shape_manual(values = c(22:25)) +
  theme_bw()
```

